---
title: "Greenhouse Gas Emissions"
author: "Tilina, Alan, Kenedi, and Shelby"
date: '2022-06-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)

# Greenhouse gas emissions by country-year from 1750 - 2020
ghg <- read_csv('https://raw.githubusercontent.com/ericmkeen/capstone/master/co2.csv')

# Greenhouse gas emissions by country-year-sector from 1990 - 2018
sectors <- read_csv('https://raw.githubusercontent.com/ericmkeen/capstone/master/co2_sectors.csv')
```

# 1. Present-Day Patterns in Co2 Emissions

## Entities
Entities with the Highest Co2 Emissions:
```{r echo = FALSE}
#Entity with highest co2
ent <- sectors %>% 
  filter(year >= 2000 
         & entity != 'World' 
         & entity != 'European Union (27)') %>% 
  group_by(entity) %>% 
  summarize(total = sum(co2, na.rm = TRUE)/1000000000) %>% 
  arrange(desc(total)) %>%
  head(10) %>% 
  arrange(total) %>% 
  mutate(entity = factor(entity, levels = entity))

options(scipen = 999)
```

```{r echo = FALSE}
ggplot(data = ent) +
  geom_col(aes(x = entity, y = total, color = entity)) +
  coord_flip() + 
  labs(y = "Total Co2 in Billions", x = "Entity", title = "Entities With Highest Co2 Emissions") + 
  scale_color_hue() +
  theme(legend.position = "none")
```

Entities with the Lowest Co2 Emissions:
```{r echo = FALSE}
#Entity with lowest co2
entLow <- sectors %>% 
  filter(year >= 2000 
         & entity != 'World' 
         & entity != 'European Union (27)') %>% 
  group_by(entity) %>% 
  summarize(total = sum(co2, na.rm = TRUE)/1000000000) %>% 
  arrange(total) %>%
  head(10) %>% 
  arrange(desc(total)) %>% 
  mutate(entity = factor(entity, levels = entity))

options(scipen = 999)
```

```{r echo = FALSE}
ggplot(data = entLow) +
  geom_col(aes(x = entity, y = total, color = entity)) +
  coord_flip() + 
  labs(y = "Total Co2 in Billions", x = "Entity", title = "Entities With Lowest Co2 Emissions") + 
  scale_color_hue() +
  theme(legend.position = "none")
```


## Sectors
Sectors with the Highest and Lowest Co2 Emissions:
```{r echo = FALSE}
#Sector with highest/lowest
sect <- sectors %>% 
  filter(year >= 2000) %>% 
  group_by(sector) %>% 
  summarize(total = sum(co2, na.rm = TRUE)/1000000000) %>% 
  arrange(total) %>%
  mutate(sector = factor(sector, levels = sector))

options(scipen = 999)
```

```{r echo = FALSE}
ggplot(data = sect) +
  geom_col(aes(x = sector, y = total, color = sector)) +
  coord_flip() + 
  labs(y = "Total Co2 in Billions", x = "Sector", title = "Economic Sectors and Their Co2 Emissions") + 
  scale_color_hue() +
  theme(legend.position = "none") 
```

## Countries
Regions with highest Emissions:
```{r}
#Rename columns
ghg1<-ghg %>% rename(country = Entity)
#Read in data
continents<-read_csv("https://github.com/alankespinoza/DataLab_2022/blob/master/BootCamp/Day8/csvData.csv?raw=true")
#Merge with continent to have country to its continent
data1<-merge(ghg1,continents, by = "country")

#Data that will give me total co2 for each continent
data1<-data1 %>% group_by(continent) %>% filter(Year >1999) %>% 
  summarize(total = sum(`Annual CO2 emissions (zero filled)`))

#Graph 
ggplot(data = data1, aes(x = continent, y = total/1000000000, fill = continent)) +
  geom_bar( stat = 'identity') +
  theme(legend.position = 'none') +
  labs(title = 'Cotinent with highest C02 Emission' ,
         subtitle = 'June 10 2022' ,
         caption = 'DataLab 2022 | Alan, Kenedi, Shelby, Tilina') +
  ylab('Total in Billions') +
  xlab('Continents')

```



# 2. Long-Term Trends in Carbon Emissions Over Time
Sector-by-Sector:
```{r echo = FALSE}
sect <- sectors %>% 
  group_by(sector, year) %>% 
  summarize(total = sum(co2, na.rm = TRUE)/1000000000) 
```

```{r echo = FALSE}
ggplot(data = sect) + 
  geom_line(aes(x = year, y = total, color = sector)) +
  labs(title = 'Long Term Carbon Emissions by Sector', 
       x = 'Year', y = 'Co2 in Billions')
```


Country-by-Country:
```{r echo = FALSE}
count <- sectors %>% 
  filter(entity == 'China' |
           entity == 'United States' |
           entity == 'India' |
           entity == 'Russia' |
           entity == 'Japan') %>% 
  group_by(entity, year) %>% 
  summarize(total = sum(co2, na.rm = TRUE)/1000000000)

ggplot(data = count) + 
  geom_line(aes(x = year, y = total, color = entity)) +
  labs(title = 'Long Term Carbon Emissions by County', 
       x = 'Year', y = 'Co2 in Billions')
```


Globally:

```{r}
# Figure out the long-term trends in CO2 emissions over year by region
data <- merge(ghg1,continents,by = 'country')
ltd <- data %>%
  group_by(continent, Year) %>%
  summarise(totalco2 = sum (`Annual CO2 emissions (zero filled)`))
# Make a graph
ggplot(data= ltd) +
  geom_line(aes(x= Year, y = totalco2/1000000000, color= continent))+
  labs(title= 'CO2 Emissions', y= 'Billions') +
  scale_x_continuous(n.breaks= 16)

# Figure out the long-term trends in CO2 emissions over year by region
ltd <- data %>%
  group_by(continent, Year) %>%
  summarise(totalco2 = sum (`Annual CO2 emissions (zero filled)`))
# Make a graph
ggplot(data= ltd) +
  geom_line(aes(x= Year, y = totalco2/1000000000, color= continent))+
  labs(title= 'CO2 Emissions', y= 'Billions') +
  scale_x_continuous(n.breaks= 16)
# Rename the entity column 'country' and merge the same countries by continent
ghg <- ghg %>%
  rename(country= Entity)
#data <- merge(ghg,continents,by = 'country')
# Group by continent from greater than 2000 and total the amount of Co2 emissions
newdata <- data %>%
  group_by(continent) %>%
  filter(Year >= 2000) %>%
  summarise(co2total = sum(`Annual CO2 emissions (zero filled)`))
# Make a graph of that^
ggplot(data = newdata) +
  geom_col(aes(x = continent, y= co2total/1000000000, fill= continent)) +
  labs(title= 'Highest vs. Lowest CO2 Emissions', subtitle= 'From 2000 - 2018', y= 'Billions', x= 'Continent')+
  theme(legend.position = 'none')
options(scipen = 999)

ggplot(data = newdata) +
  geom_col(aes(x = continent, y= co2total/1000000000, fill= continent)) +
  labs(title= 'Highest vs. Lowest CO2 Emissions', subtitle= 'From 2000 - 2020', y= 'Billions', x= 'Continent')+
  theme(legend.position = 'none')
options(scipen = 999)
```


# 3. Details on These Trends in the Top 5 Most-Emitting Countries

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(leaflet)
library(ggplot2)
library(ggthemes)
library(flexdashboard)
library(knitr)
library(plotly)
library(readr)

#install.packages('plotly')

sectors1<-sectors %>% rename(country = entity)
#Find top 10 countries
top10<-sectors1 %>% group_by(country) %>% filter(year >1999) %>% 
  summarize(total = sum(co2, na.rm = TRUE)) %>% arrange(desc(total)) %>% head(10)
#Find total for each sector in top 10 countries
data2<-sectors1 %>% filter( country %in% top10$country) %>% group_by(country,sector) %>%
  filter(year >1999) %>% summarize(total = sum(co2, na.rm = TRUE))
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
#graph
g<-ggplot(data = data2, aes(x = country, y = total/1000000000, fill = sector)) +
  geom_bar( stat = 'identity') +
  labs(title = 'Top 10 Countries with highest C02 Emission' ,
         subtitle = 'June 10 2022' ,
         caption = 'DataLab 2022 | Alan, Kenedi, Shelby, Tilina') +
  ylab('Total in Billions') +
  xlab('Continents') +
  theme(legend.position = 'none')
#interactive!
ggplotly(g)
```



